---
title: "Project2DATA607DataSet2"
author: "Banu Boopalan"
date: "11/4/2019"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
  ioslides_presentation: default
  pdf_document: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(tidyr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(data.table)
library(readr)
library(knitr)
#Install KableExtra package and use that
library(kableExtra)
```

## Project 2 :DataSet2 Analysis. This dataset has Year data spread in a wide format. Also it has some special characters and formats to be cleaned up.

Your task is to: For each of the three chosen datasets: • Create a .CSV file (or optionally, a MySQL database!) that includes all of the information included in the dataset.  You’re encouraged to use a “wide” structure similar to how the information appears in the discussion item, so that you can practice tidying and transformations as described below. 

DATASET SOURCE: World Health Organization Data Repository
URL link: http://apps.who.int/gho/data/node.main.618?lang=en						


```{r}
myurl <- "https://raw.githubusercontent.com/BanuB/DATA607Project2/master/HIVDatasetSourceWorldHealthOrganizationData.csv"
csvdata3  <- read.csv(file=myurl, header=TRUE,sep=",",stringsAsFactors = FALSE,na.strings=c("NA"))
filename<- "HIVWHODataset2.csv"
```

## After reading in data, rename columns


```{r}
str(csvdata3)
view(csvdata3)

names(csvdata3)[2] <- "EstantiretrocoveragepeopleHIV2018"
names(csvdata3)[3] <- "ReportedantiretropeopleHIV2018"
names(csvdata3)[4] <- "EstallHIV2018"
names(csvdata3)[5] <- "EstallHIV2010"
names(csvdata3)[6] <- "EstallHIV2005"
names(csvdata3)[7] <- "EstallHIV2000"
csvdata3 %>% kable() %>% kable_styling()

```


## apply gsub to cleanup strings

```{r}
raw.table1 <- csvdata3 %>%
  mutate(EstantiretrocoveragepeopleHIV2018 = type.convert(str_extract(EstantiretrocoveragepeopleHIV2018, "^[0-9]+"),na.strings = "NA", as.is = FALSE, dec= ".")
#         ,ReportedantiretropeopleHIV2018 = type.convert(str_extract(ReportedantiretropeopleHIV2018, "^[0-9]+"),na.strings = "NA", as.is = FALSE, dec= ".")
,ReportedantiretropeopleHIV2018 = type.convert(str_extract(gsub("\\s+", "", ReportedantiretropeopleHIV2018), "^[0-9]+") ,na.strings = "NA", as.is = FALSE, dec= ".")
         ,EstallHIV2018 = type.convert(gsub("\\[","", str_extract(gsub("\\s+", "",EstallHIV2018), "[0-9].*\\[")),na.strings = "NA", as.is = FALSE, dec= ".")
         ,EstallHIV2010 = type.convert(gsub("\\[","", str_extract(gsub("\\s+", "",EstallHIV2010), "[0-9].*\\[")),na.strings = "NA", as.is = FALSE, dec= ".")
         ,EstallHIV2005 = type.convert(gsub("\\[","", str_extract(gsub("\\s+", "",EstallHIV2005), "[0-9].*\\[")),na.strings = "NA", as.is = FALSE, dec= ".")
         ,EstallHIV2000 = type.convert(gsub("\\[","", str_extract(gsub("\\s+", "",EstallHIV2000), "[0-9].*\\[")),na.strings = "NA", as.is = FALSE, dec= ".")
         
  )
#raw.table1[3] <- lapply(raw.table1[3], as.numeric)
str(raw.table1)
raw.table1 %>% kable() %>% kable_styling()
```

## arrange data and order


```{r}
raw.table3 <- raw.table1 %>%
  select(Country,EstantiretrocoveragepeopleHIV2018,ReportedantiretropeopleHIV2018) %>% arrange(desc(ReportedantiretropeopleHIV2018))


data_long3 <-  subset (raw.table3,raw.table3$EstantiretrocoveragepeopleHIV2018 >60) 
data_long4 <-  subset (raw.table3,raw.table3$EstantiretrocoveragepeopleHIV2018 <50) 

str(data_long3)
data_long3<- data_long3 %>% arrange(desc(ReportedantiretropeopleHIV2018))
data_long4<- data_long4 %>% arrange(desc(ReportedantiretropeopleHIV2018))


data_long3 %>% kable() %>% kable_styling()
data_long4 %>% kable() %>% kable_styling()



```

## Plot


```{r}
data_long2 <- gather(data_long3, Year, HIVtotal, ReportedantiretropeopleHIV2018)
data_long2a <- gather(data_long4, Year, HIVtotal, ReportedantiretropeopleHIV2018)
data_long2c <- gather(raw.table1 , Year, HIVtotal, 4:7) 
#data_long2d <- head(data_long2c, n=1000)
#data_long2c  %>% kable() %>% kable_styling()

ggplot(data_long2, aes(x=Country, y=HIVtotal, colour = Year, group = Year, fill = EstantiretrocoveragepeopleHIV2018)) + geom_line(linetype = "dashed") + geom_point(shape = 22, size = 3, fill = "white")+ ggtitle("Countries (% of reported people receiving therapy > 60%), reported number living with HIV") + labs(x="Country", y="HIVTotal")+
  theme(axis.text.x = element_text(angle=90))
ggplot(data_long2a, aes(x=Country, y=HIVtotal, colour = Year, group = Year, fill = EstantiretrocoveragepeopleHIV2018)) + geom_line(linetype = "dashed") + geom_point(shape = 22, size = 3, fill = "white")+ ggtitle("Countries (% of reported people receiving therapy < 50%) , reported number living with HIV") + labs(x="Country", y="HIVTotal")+
  theme(axis.text.x = element_text(angle=90))

data_long2d <-  subset(data_long2c, HIVtotal > 500000 ) 
data_long2d  %>% kable() %>% kable_styling()
data_long2e <-  subset(data_long2c, Country %in% c('Estonia','Gambia','Japan')) 
data_long2e  %>% kable() %>% kable_styling()

g1 <- ggplot(data_long2e, aes(x=Country, y=HIVtotal, group = Year,fill=Year))
g1 + geom_bar(stat="identity", width = 2) + 
  theme(axis.text.x = element_text(angle=90, vjust=1)) + 
  labs(title="HIV total", 
       subtitle="",
       x="Country",
       y="HIVtotal") +   facet_wrap(~ Year)


# Plot
g <- ggplot(data_long2d, aes(Country,HIVtotal ))
g + geom_bar(stat="identity", width = 1, fill="Red") + 
      labs(title="Bar Chart", x= "Country", y= "HIVtotal") + 
          theme(axis.text.x = element_text(angle=65, vjust=1.0))
```


