---
title: "Project2DATA607DataSet3"
author: "Banu Boopalan"
date: "11/4/2019"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
  ioslides_presentation: default
  pdf_document: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(tidyr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(data.table)
library(readr)
library(knitr)
#Install KableExtra package and use that
library(kableExtra)
```

## Project 2 :DataSet3 Analysis. This dataset has Year data spread in a wide format. So, we can transform and create a long format.

Your task is to: For each of the three chosen datasets: • Create a .CSV file (or optionally, a MySQL database!) that includes all of the information included in the dataset.  You’re encouraged to use a “wide” structure similar to how the information appears in the discussion item, so that you can practice tidying and transformations as described below. 

DATASET SOURCE: UN World Populalation Dataset
Suggested citation: United Nations, Department of Economic and Social Affairs, Population Division (2019). World Population Prospects 2019, Online Edition. Rev. 1.							


```{r}
myurl <- "https://raw.githubusercontent.com/BanuB/DATA607Project2/master/WorldTOTAL_POPULATION_BOTH_SEXESFinal.csv"
csvdata4 <- read.csv(file=myurl, header=TRUE,sep=",",stringsAsFactors = FALSE,na.strings=c("NA"))
filename<- "WorldPopulationDataset3.csv"
```

## After reading in data, keep only some records and certain coloums using subset


```{r}
mydataset <- subset(csvdata4, csvdata4$X.4 == 'Country/Area' |csvdata4$X.4 == 'Type' )
str(mydataset)
head(mydataset, n=10)

mydataset1 <- mydataset %>%
  select(X.1,X.55:X.75)
str(mydataset1)
head(mydataset1)
```


## Show one variable Y2020 only

```{r}
mydataset1[22]
```

## Rename columns and keep only 2000-2020 for reporting and drop the first row to keep only values we need


```{r}
names(mydataset1)[1] <- "Country"
names(mydataset1)[2] <- "Y2000"
names(mydataset1)[3] <- "Y2001"
names(mydataset1)[4] <- "Y2002"
names(mydataset1)[5] <- "Y2003"
names(mydataset1)[6] <- "Y2004"
names(mydataset1)[7]<- "Y2005"
names(mydataset1)[8] <- "Y2006"
names(mydataset1)[9] <- "Y2007"
names(mydataset1)[10] <- "Y2008"
names(mydataset1)[11] <- "Y2009"
names(mydataset1)[12] <- "Y2010"
names(mydataset1)[13] <- "Y2011"
names(mydataset1)[14] <- "Y2012"
names(mydataset1)[15] <- "Y2013"
names(mydataset1)[16] <- "Y2014"
names(mydataset1)[17] <- "Y2015"
names(mydataset1)[18] <- "Y2016"
names(mydataset1)[19] <- "Y2017"
names(mydataset1)[20] <-"Y2018"
names(mydataset1)[21] <- "Y2019"
names(mydataset1)[22] <- "Y2020"

mydataset2 <- mydataset1  %>% filter(!str_detect(Country, 'Region'))
str(mydataset2)

```

## Run function Check to remove space and convert to numeric


```{r}
check<-function(x){
  return(as.numeric(str_replace_all(x, "\\s+", "")))
}
mydataset2[2:22] <- apply(mydataset2[2:22], MARGIN=2, FUN=check)
str(mydataset2)
```

## Gather and Arrange long dataset


```{r}
WPmydataset<- mydataset2 %>% gather("Year", "Population", Y2000:Y2020)
WPmydataset %>% arrange(Country,Year,Population) 
head(WPmydataset)
```

## Filter only top 10 countries for Y2020


```{r}
WP1 <- arrange(WPmydataset, Year)
WP10 <- WP1  %>% filter(Year == 'Y2020')
WP2 <- WP10 %>% arrange(desc(Population)) 
WP3 <- head(WP2, n=10)

WP3 %>% kable() %>% kable_styling()
```

## Plot population using ggplot for top countries


```{r}
ggplot(WP3, aes(x = reorder(Country, -Population), y = Population, fill = Country)) +
  geom_bar(stat = "identity") +
    xlab("Country") +
  ylab("Population") +
  ggtitle("Top Global Population (in thousands)") +
  theme(axis.text.x = element_text(angle = 45,hjust = 1)) + 
  theme(plot.title = element_text(hjust = 0.5))

```
