---
title: "Project2DATA607DataSet1"
author: "Banu Boopalan"
date: "11/4/2019"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
  ioslides_presentation: default
  pdf_document: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(tidyr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(data.table)
library(readr)
library(knitr)
#Install KableExtra package and use that
library(kableExtra)
library(reshape2)

```

## Project 2 :DataSet1 Analysis. This dataset has Year data spread in a wide format. Also it has some special characters and formats to be cleaned up.

Your task is to: For each of the three chosen datasets: • Create a .CSV file (or optionally, a MySQL database!) that includes all of the information included in the dataset.  You’re encouraged to use a “wide” structure similar to how the information appears in the discussion item, so that you can practice tidying and transformations as described below. 

DATASET SOURCE: US Census Data
URL link: https://www.census.gov/data/tables/time-series/demo/income-poverty/cps-pinc/pinc-06.2018.html						


```{r}
myurl <- "https://raw.githubusercontent.com/BanuB/DATA607Project2/master/census2femalecsv1.csv"
csvdata2  <- read.csv(file=myurl, header=TRUE,sep=",",stringsAsFactors = FALSE,na.strings=c("NA"))
filename<- "CensusFemaleDataset1.csv"
```

## After reading in data, view file


```{r}
str(csvdata2)
raw.tablefemale <- data.frame(csvdata2)
names(raw.tablefemale)[1]<-"Occupation"
str(raw.tablefemale)
```


## tidy data and rename columns

```{r}
raw.tablefemale1 <- raw.tablefemale %>%
  mutate(Occupation = gsub("[[:punct:]]", "", Occupation)) 

raw.tablefemale2 <- raw.tablefemale1 %>%
  mutate(Occupation = gsub("[^a-zA-Z]","", Occupation))
raw.tablefemale2  %>% kable() %>% kable_styling()


```

## arrange data and order


```{r}
melt.tablefemale <- melt(raw.tablefemale2,id.vars="Occupation",variable.name = "Earnings", value.name = "Count")
melt.tablefemale3 <- subset (melt.tablefemale, Count != "(B)")
#head(melt.tablefemale,n=200)
str(melt.tablefemale3)
head(melt.tablefemale3,n=200) %>% kable() %>% kable_styling()

check<-function(x){
  return(as.numeric(gsub("[[:punct:]]","",x)))
    #gsub("[[:punct:]]", "",as.character(x))))
}
melt.tablefemale3[3] <- lapply(melt.tablefemale3[3],FUN=check)
str(melt.tablefemale3)
head(melt.tablefemale3,n=200) %>% kable() %>% kable_styling()
#melt.tablefemale3$Count
```

## filter by few occupations


```{r }
melt.tablefemale2 <- melt.tablefemale3 %>% filter(Earnings == "Median.earnings" | Earnings == "Mean.earnings" ) %>%  arrange(Earnings,desc(Count))
myselect <- c('Chief','Doctors','Nurses','FirefightersandPolice')
melt.tablefemale4 <- melt.tablefemale2   %>% filter(str_detect(Occupation, myselect) == TRUE)
head(melt.tablefemale2,n=200) %>% kable() %>% kable_styling()
melt.tablefemale4  %>% kable() %>% kable_styling() 



wptable4 <- head(melt.tablefemale4, n=30)


```


